UKB Models

### 2023-10-17 proof-of-concept_mods.Rmd
switch to modeling numeric sexMale for drop1() functionality

```{r}
ukb_df_totals <- read_csv(file="data/ukb_CN_data.csv")
ukb_df_age.filt <- read_csv(file="data/ukb_CN_data_agefilt.csv")

@@

```{r}
read_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/ukb_basic/gamlss_objs" #path to bfp models
```

@@

>>>>

```{r}
ukb_df_totals <- read_csv(file="data/ukb_CN_data.csv")
ukb_df_age.filt <- read_csv(file="data/ukb_to_model/ukb_CN_data_simsites.csv")

@@

```{r}
read_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/ukb_basic/" #path to bfp models
```

@@

Troubleshooting drop1

```{r}
model.test <- readRDS("~/Desktop/BGD_Repos/combat_biovar/ukb_basic/lh_SA_superiortemporal_ukb-CN-data-simsites_mod.rds")
drop1(model.test, what = "sigma")
model.test.drop <- gamlss(formula = lh_SA_superiortemporal ~ poly(age_days, 3) + sex,
                       sigma.formula = ~ poly(age_days, 3),
                       nu.formula = ~ 1,
                       family = BCCG, data=ukb_df_age.filt, trace = FALSE)
drop1(model.test.drop, what = "sigma")

ukb_df_age.filt <- ukb_df_age.filt %>%
  mutate(sexMale = as.numeric(case_when(sex == "Male" ~ 1,
                                        sex == "Female" ~ 0,
                                        TRUE ~ NA)))
model.test.refit <- gamlss(formula = lh_SA_superiortemporal ~ poly(age_days, 3) + sexMale,
                       sigma.formula = ~ poly(age_days, 3) + sexMale,
                       nu.formula = ~ 1,
                       family = BCCG, data=ukb_df_age.filt, trace = FALSE)
drop1(model.test.refit, what = "sigma", scope = "sexMale")
```

### 2023-10-17 ukb_combat.Rmd
troubleshooting and updating filenaming
switch to modeling numeric sexMale for drop1() functionality

```{r}
ukb_df_sites <- read_csv(file="data/ukb_CN_data_simsites.csv")
pheno_list <- readRDS(file="R_scripts/pheno_list.rds")
```

@@

#covariates - using model.matrix to dummycode sex as numeric
covar.mat <- model.matrix( ~ age_days + sex, data=ukb_df_sites)

@@

write out
```{r}
for (i in names(combat.df.list)){
  df <- combat.df.list[[i]] %>%
  write_csv(df, file=paste0("data/ukb_to_model/ukb_", as.character(i), "_data.csv"))
}
```

>>>>

```{r}
ukb_df_sites <- read_csv(file="data/ukb_to_model/ukb_CN_data_simsites.csv")
pheno_list <- readRDS(file="R_scripts/pheno_list.rds")
```

@@

# #covariates - using model.matrix to dummycode sex as numeric - 
# covar.mat <- model.matrix( ~ age_days + sexMale, data=ukb_df_sites) 
#testing with df now that mods are being fit with sexMale
covar.mat <- ukb_df_sites %>%
  dplyr::select(age_days, sexMale)

@@

write out
```{r}
for (i in names(combat.df.list)){
  df <- combat.df.list[[i]]
  write_csv(df, file=paste0("data/ukb_to_model/ukb_", i, "_data.csv"))
}
```

### 2023-10-17 ukb_curation-eda.Rmd
switch to modeling numeric sexMale for drop1() functionality

```{r}
ukb_df <- cn_data %>%
  dplyr::filter(study == "UKB") %>%
  mutate(age_yrs = round(age_days/365.25),
         log_age = log(age_days, base=10),
         TBV=(sGMV+GMV+WMV+Ventricles),
         fs_version = as.factor(fs_version),
         sex = as.factor(sex),
         sex_int = as.numeric(case_when(sex == "Male" ~ 0,
                                        sex == "Female" ~ 1)),
         sex.age = sex_int * age_days) %>%
  na.omit() #1 subj has 1 missing val

>>>>

```{r}
ukb_df <- cn_data %>%
  dplyr::filter(study == "UKB") %>%
  mutate(age_yrs = round(age_days/365.25),
         log_age = log(age_days, base=10),
         TBV=(sGMV+GMV+WMV+Ventricles),
         fs_version = as.factor(fs_version),
         sex = as.factor(sex),
         sexMale = as.numeric(case_when(sex == "Male" ~ 1,
                                        sex == "Female" ~ 0,
                                        TRUE ~ NA)),
         sex.age = sexMale * age_days) %>%
  na.omit() #1 subj has 1 missing val
